---
title: "Practical Machine Leaning Project"
output: html_document
---
###Introduction to the Project

This project looks at personal activity data with the goal of predicting
*how well* an activity is done, not just *how much* of an activity is done.
For this project a prediction model will be constructed to predict the 
*effectiveness* of activity given the *amount* of movement.

The data that will be used in this model is based on barbell lifts performed
both correctly and incorrectly in five different ways. Information about the
instructions given to the participants can be seen 
[here](http://groupware.les.inf.puc-rio.br/har).

Within the data set, there are five classes of workout results. The 
[training data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)
will be used to create a model that will be able to predict the class based on
the amount of movement during the workout.

###Data Loading

Begin by loading the both the training and the test data.
```{r}
training <- read.csv("C:\\Users\\lisar\\Documents\\R\\PML\\pml-training.csv")
testing <- read.csv("C:\\Users\\lisar\\Documents\\R\\PML\\pml-testing.csv")

# these are not being used to illinate data for faster run time
training$cvtd_timestamp <- NULL
training$user_name <- NULL
training$raw_timestamp_part_1 <- NULL
training$raw_timestamp_part_2 <- NULL
training$num_window <- NULL
for (col_name in names(training)){
        if (col_name %in% c("user_name", "classe")){
                training[ , col_name] <- as.factor(training[ , col_name])
        } else {
                training[ , col_name] <- as.numeric(training[ , col_name])
                if (nrow(training[is.na(training[ , col_name]) ,])){
                        training[ ,col_name] <- NULL
                }
        }
}

# select the column names to use for prediction model
cols <- c("X" , "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
             "new_window", "num_window", "cvtd_timestamp")
colsneeded <- setdiff(names(training), cols)
newtraining <- subset(training, T, colsneeded)
```

###Data Modeling

Now that there is a training set of data, fit the model. For the data that is 
provided, a random forest model is an accurate model type, so that will be used.

```{r eval=FALSE}
library(caret)
set.seed(2)

# partition data for modeling
trainingData <- createDataPartition(y=newtraining$classe, p=0.7, list=F)
newtraining2 <- newtraining[trainingData,]
testing2 <- newtraining[-trainingData,]

# using the random forest model to predict classe results
testFit <-train(classe ~ ., data=newtraining2, method="rf")
```

Now create a confusion matrix to show that the model's accuracy is acceptable.

```{r eval=FALSE}
confusionMatrix(predict(testFit), newtraintraining2$classe)
```

### Validate the Random Forest Model

Test the model on the test dataset.

```{r eval=FALSE}
confusionMatrix(predict(testFit), testing2$classe)
```

This demonstrates that the model is accurate and acceptable for the given datasets.

### Summary of results

Based on the results above, the random forest model is a good selection for
this predictive model. It produced a 99%+ accuracy rate. The model created for
this project is good predictor of workout efficiency given the amount of 
movement performed.
